MS Technical Paper: \\ Placement Algorithms for Heterogeneous FPGAs

### 2025-09-05 latex/paper/sections/02-abstract.tex
readme graphics

\section{Abstract}
fdsafdsafdsa. \\ 
fdsafdsafdsa. \\ 
fdsafdsafdsa. \\ 
fdsafdsafdsa. \\ 
fdsafdsafdsa. \\ 
fdsafdsafdsa. \\ 
fdsafdsafdsa. \\ 
fdsafdsafdsa. \\ 
fdsafdsafdsa. \\ 
fdsafdsafdsa. \\ 
fdsafdsafdsa. \\ 
fdsafdsafdsa. \\ 
fdsafdsafdsa. \\

>>>>

\section{Abstract}
Placement remains one of the most computationally demanding stages of FPGA computer-aided design (CAD) flows, directly influencing routability, timing, and design productivity. 
This paper presents a modular placement framework targeting Xilinx 7-Series devices, implemented using the RapidWright API. 
The flow first clusters multi-cell structures—such as CARRY4 chains, DSP48E1 cascades, and LUT–FF pairs—into prepacked groups, then maps them to physical sites before applying a simulated annealing (SA) placer. 
The SA engine minimizes half-perimeter wirelength (HPWL) using a geometric cooling schedule, with support for both random and directed moves. 
The result is a legal, wirelength-optimized placement that forms a foundation for experimenting with alternative cost functions and analytical or hybrid placement strategies.

### 2025-09-05 latex/paper/sections/03-introduction.tex
readme graphics

For these reasons, FPGA placement remains a critical research effort even today. 
In this paper, we study and implement established placement methods. 
To do this, we use the RapidWright API, which is a semi-open-source research effort from AMD/Xilinx that enables custom solutions to FPGA design implementations and design tools that are not offered by their industry-standard FPGA environment, Vivado. 
We implement multiple variations of simulated annealing placers for Xilinx's 7-series FPGAs, with an emphasis on minimizing total wirelength while mitigating runtime. 
Our implementation is organized into three consecutive substages. 
The \textbf{prepacking} stage involves traversing a raw EDIF netlist to identify recurring cell patterns—such as CARRY chains, DSP cascades, and LUT-FF pairs—that are critical for efficient mapping and legalization. 
In the subsequent \textbf{packing} stage, these identified patterns, along with any remaining loose cells, are consolidated into SiteInst objects that encapsulate the FPGA’s discrete resource constraints and architectural nuances. 
Finally, the \textbf{placement} stage employs a simulated annealing (SA) algorithm to optimally assign SiteInst objects to physical sites, aiming to minimize total wirelength while adhering to the constraints of the 7-series architecture.

>>>>

For these reasons, FPGA placement remains a critical research effort even today. 
In this paper, we study and implement established placement methods. 
To do this, we use the RapidWright API, which is a semi-open-source research effort from AMD/Xilinx that enables custom solutions to FPGA design implementations and design tools that are not offered by their industry-standard FPGA environment, Vivado. 
We implement multiple variations of simulated annealing placers for Xilinx's 7-series FPGAs, with an emphasis on minimizing total wirelength while mitigating runtime. 
Our implementation is organized into three consecutive substages. 
The \textbf{prepacking} stage involves traversing a raw EDIF netlist to identify recurring cell patterns—such as CARRY chains, DSP cascades, and LUT-FF pairs—that are critical for efficient mapping and legalization. 
In the subsequent \textbf{packing} stage, these identified patterns, along with any remaining loose cells, are consolidated into \texttt{SiteInst} objects that encapsulate the FPGA’s discrete resource constraints and architectural nuances. 
Finally, the \textbf{placement} stage employs a simulated annealing (SA) algorithm to optimally assign \texttt{SiteInst} objects to physical \texttt{Sites}, aiming to minimize total wirelength while respecting the constraints of the 7-series architecture.

### 2025-09-05 latex/paper/sections/12-future_work.tex
readme graphics

The connectivity matrix $\boldsymbol{Q_x}$ encodes 2-pin connections between movable instances while $\boldsymbol{c_x}$ encodes 2-pin connections between movable instances and fixed instances   
Movable instances are for example \texttt{SLICEL}s, DSP, and RAM, while fixed instances can be IO pads and clock buffers.
We distinguish fixed instances because their placement cannot be optimized and can be thought of as constraints encoded into the objective.
For a design with $M$ pins belonging to movable instances and $F$ pins belonging to fixed instances, $\boldsymbol{Q}_x \in {\mathbb{R}}^{M \times M}$ and $\boldsymbol{c}_x \in {\mathbb{R}}^F$.

>>>>

The connectivity matrix $\boldsymbol{Q_x}$ encodes 2-pin connections between movable instances while $\boldsymbol{c_x}$ encodes 2-pin connections between movable instances and fixed instances   
Movable instances are for example \texttt{SLICEL}s, DSP, and RAM, while fixed instances can be IO pads and clock buffers.
We distinguish fixed instances because their placement cannot be optimized and can be thought of as constraints encoded into the objective.
For a design with $M$ pins belonging to movable instances and $F$ pins belonging to fixed instances, $\boldsymbol{Q}_x \in {\mathbb{R}}^{M \times M}$ and $\boldsymbol{c}_x \in {\mathbb{R}}^F$.
We can construct $\boldsymbol{Q}_x$ and $\boldsymbol{c}_x$ formally by taking the graph Laplacian matrix $L_x$ of our netlist reduced into a graph and partitioning the matrix to separate M-M from M-F weights to form $\boldsymbol{Q}_x$ and $\boldsymbol{c}_x$.
However, we can simply follow a shortcut outlined in SimPL \cite{SimPL}.
If we take our netlist reduced into a graph
